---
title: "First Markdown"
author: "Naylor"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
data(mtcars)
library(tidyverse)
library(modelr)
```

Here's a summary of mtcars. 
```{r}
summary(mtcars)
```



Here's a plot of miles per gallon as a function of hp
```{r, message=FALSE, warning=FALSE, fig.align='center', fig.width=3, fig.height=4}
ggplot(mtcars, aes(x=hp,y=mpg))+
  geom_point()+
  geom_smooth(method="lm")
```


What about text??

# Text
## Text
### Text

**bold**
*italic*

We would like to model the effect of various factors on miles per gallon.
Let's start with a simple model where horsepower is the only predictor:

```{r, echo=FALSE}
mod1 <- glm(dat=mtcars,
            formula = mpg~hp)
mod2 <- glm(data = mtcars,
            formula = mpg ~ poly(hp,2))
summary(mod1)
summary(mod2)
```

We built two versions of the model. The first is linear, the second is second-order
polynomial...both using hp as the only predictor.

Let's make a plot:
```{r, echo=FALSE}

gather_predictions(mtcars,mod1,mod2) %>%
  ggplot(aes(x=hp,y=mpg))+
  geom_point(aes(x=hp,y=mpg))+
  geom_point(size=3,alpha=.25)+
  geom_point(aes(y=pred,color=model))+
  geom_smooth(aes(y=pred,color=model))
```

Which model is the best fit?
Let's look at the root mean=sp of error:

```{r}
mean(residuals(mod1)^2)
mean(residuals(mod2)^2)
```


Here, we are fitting a curved line using natural splines with  3rd order polynomial function..
```{r}
library(splines)
ggplot(data=mtcars, aes(x=hp,y=mpg))+
  geom_smooth(method="lm", formula = y ~ ns(x,3))

```


<br/><br/><br/><br/><br/><br/><br/><br/><br/><br/><br/>

[hyperlink](https://www.google.com/) 

some more text
