---
title: "Assignment_9"
author: "Bowen"
date: "10/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r admissions, message=FALSE, warning=FALSE, echo=FALSE}
df_admissions <- read.csv("../../Data/GradSchool_Admissions.csv", stringsAsFactors = FALSE)
data(df_admissions)
library(tidyverse)
library(modelr)
library(GGally)
```


### Below is a summary for admission rates
```{r admission2, message=FALSE, echo=FALSE, fig.align='center'}
summary(df_admissions)
```


### Below is a plot for predicted admissions based on GPA:
```{r admissions2, message=FALSE, fig.align='center', echo=FALSE}
df_admissions$admit <- as.logical(df_admissions$admit)
df_admissions$rank <- factor(df_admissions$rank)
mod1 <- glm(data = df_admissions, formula = admit ~ gre + gpa + rank, family = "binomial")
df_withpredictions <- add_predictions(df_admissions, mod1, type = "response")
ggplot(df_withpredictions, aes(x = gpa, y= pred, color= as.factor(rank)))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(x = "GPA", y= "Prediction", title = "(p1) Grad School Admissions Predictions Based on GPA", subtitle = "Including University level ranking")
```


### The next two graphs show prediction percentages of a student with a 3.5 GPA from a rank 1 school vs a rank 2 school given different GRE scores:
```{r hypothetical gre, message=FALSE, echo=FALSE, fig.align='center'}
hypothetical_data <- data.frame(gre=c(500, 550, 600, 650, 700, 750),
                                gpa=c(3.5, 3.5, 3.5, 3.5, 3.5, 3.5),
                                rank=c("1", "1", "1", "1", "1", "1"))
df2_withpredictions <- add_predictions(hypothetical_data, mod1, type = "response")
ggplot(df2_withpredictions, aes(x = gre, y= pred))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(x = "GRE Score", y= "Prediction", title = "(p2) Grad School Admissions Rate based on GRE", subtitle = "From a rank 1 University")
```








```{r plot, message=FALSE, echo=FALSE, fig.align='center'}
hypothetical_data2 <- data.frame(gre=c(500, 550, 600, 650, 700, 750),
                                gpa=c(3.5, 3.5, 3.5, 3.5, 3.5, 3.5),
                                rank=c("2", "2", "2", "2", "2", "2"))
df3_withpredictions <- add_predictions(hypothetical_data2, mod1, type = "response")
ggplot(df3_withpredictions, aes(x = gre, y= pred))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(x = "GRE Score", y= "Prediction", title = "(p3) Grad School Admissions Rate based on GRE", subtitle = "From a rank 2 University")
```



We can see that there is a significant difference in predicted grad school admissions from those coming from a rank 1 school vs rank 2.


### Below is a graph that we can compare to the one above. Everything is the same except included this time is admit as a function of GRE and GPA together: 
```{r plot4, echo=FALSE, message=FALSE, fig.align='center'}
mod3 <- glm(data = df_admissions, formula = admit ~ gre * gpa + rank, family = "binomial")
hypothetical_data3 <- data.frame(gre=c(500, 550, 600, 650, 700, 750),
                                gpa=c(3.5, 3.5, 3.5, 3.5, 3.5, 3.5),
                                rank=c("2", "2", "2", "2", "2", "2"))
df4_withpredictions <- add_predictions(df_admissions, mod3, type = "response")
ggplot(df4_withpredictions, aes(x = gre, y= pred, color = rank))+
  geom_point()+
  geom_smooth(method = "lm", color = "red")+
  labs(title = "(p4)")
```


This is using the original/raw data, without predictions. This plot in particular is not very helpful...
```{r plot5, message=FALSE, echo=FALSE, fig.align='center'}
ggplot(df_admissions, aes(x = admit, y= gpa, color = as.factor(rank)))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "(p5)")
```





### This plot only takes school ranking in to account. This is a prediction of 4 students with the same GPA and GRE but coming from differently ranked schools. It is interesting to note how big of a difference the school rank alone changes things. 

```{r plot6, message=FALSE, echo=FALSE, fig.align='center'}
hypothetical_data4 <- data.frame(gre=c(650, 650, 650, 650),
                                gpa=c(3.5, 3.5, 3.5, 3.5),
                                rank=c("1", "2", "3", "4"))
df6_withpredictions <- add_predictions(hypothetical_data4, mod1, type = "response")
ggplot(df6_withpredictions, aes(x = rank, y= pred))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(x = "School rank", y= "Prediction", title = "(p6) Grad School Admissions Rate Predictions based on school ranking", subtitle = "All GRE and GPA scores are the same")
```




